This is loading and analyzing the data to make PCA of the annotated genes

Loading in packages and making a collected dataframe, where the duplicates within the clusters have been removed
```{r}
library('dplyr')
library('tidyr')
library('tidyverse')
library('ggplot2')
library('devtools')
library('factoextra')
library("ggfortify")
library('pcaMethods')
library('ggrepel')
```


```{r}
GeneClustering <- read.csv("New order uTSS mm10 quant FPKM  - 1kb quantile corrected 10c.txt", sep = "\t")
GeneAnnotations <- read.csv("dbtmee_cluster_gene_v2.tsv", sep = "\t")
GeneAnnotations <- as.data.frame(GeneAnnotations)

#removing duplicates from within the clusters
GeneClustering_filtered <- GeneClustering %>% distinct(Group.order, Gene.Name, .keep_all = TRUE)
names(GeneAnnotations)[names(GeneAnnotations) == 'Gene'] <- 'Gene.Name'


head(GeneAnnotations)
head(GeneClustering_filtered)
```


Here there are around 10k genes that are filtered out as there is no annotation for them
```{r}
mergedDF <- merge(GeneClustering_filtered, GeneAnnotations, by = "Gene.Name", all = FALSE)
#mergedDF <- merge(GeneClustering_filtered, GeneAnnotations, by = "Gene.Name", all = TRUE)
```

Making a new dataframe only keeping the columns needed and then renaming them to fit the data
(classification, cluster, group1, group2, group3, etc.)
```{r}
# df <- mergedDF %>% select("Cluster.y", "Order.based.on.stages", "Gene.Name", "X6.p7.Az.250.16_S14_R1_001", "X4.p10B.AZ.250.14_S10_R1_001", "X1.p12.AZ.200_S1_L001_R1_001_pooled", "X3.NSN.AZ.200_S3_L001_R1_001_pooled", "X4.SN.AZ.200_S4_L001_R1_001_pooled", "X5.MII.AZ.300_S5_L001_R1_001_pooled", "X8.8c.AZ.460_S8_L001_R1_001", "X11.mor.AZ460_S11_L001_R1_001", "X13.bla.AZ460_S13_L001_R1_001")

 df <- mergedDF %>% select("Cluster.y", "Order.based.on.stages", "Gene.Name", "X6.p7.Az.250.16_S14_R1_001.quantile.normalised", "X4.p10B.AZ.250.14_S10_R1_001.quantile.normalised", "X1.p12.AZ.200_S1_L001_R1_001_pooled.quantile.normalised", "X3.NSN.AZ.200_S3_L001_R1_001_pooled.quantile.normalised", "X4.SN.AZ.200_S4_L001_R1_001_pooled.quantile.normalised", "X5.MII.AZ.300_S5_L001_R1_001_pooled.quantile.normalised", "X8.8c.AZ.460_S8_L001_R1_001.quantile.normalised", "X11.mor.AZ460_S11_L001_R1_001.quantile.normalised", "X13.bla.AZ460_S13_L001_R1_001.quantile.normalised")


df <- setNames(df, c("Gene_type","Cluster","Gene_name", "P7", "P10", "P12", "NSN", "SN", "MII", "8cell", "Morula", "Blastocyst"))

#names(mergedDF)
```


```{r}
data <- df[,4:12]

numerical_data <- apply(data, 2, function(x) as.numeric(gsub(",", ".", x)))
```

Using bioconductor:
```{r}
#pcoutput <- pca(t(numerical_data), nPcs = 10, scale = "uv", center = TRUE, cv ="none")

pcoutput <- pca(t(numerical_data))
dat <- as.data.frame(scores(pcoutput)[,1:2])
dat$celltype <- rownames(dat)


p <- ggplot(dat, aes(PC1, PC2, colour=celltype, label=celltype)) + 
  geom_point(size=3) + 
  geom_text_repel() + 
  theme_minimal() +
  xlab("PC1 39.6%") + 
  ylab("PC2 21.9%")
p


```

```{r}
pcoutput
```







Using prinComp:
```{r}
data_normalized <- scale(numerical_data)

data.pca <- princomp(numerical_data)

```

```{r}
autoplot(data.pca)
# Graph of the variables
fviz_pca_var(data.pca, col.var = "black")
fviz_pca_ind(data.pca)
```















