---
title: "OptimalClusterNumber"
output: html_document
date: "2024-07-27"
---

Loading packages
```{r}
library('ggplot2')
library('tidyverse')
```


Estimating the optimal cluster number from one single heatmap
```{r}
# Number of bins definec
pos <- -99:99

folder <- 'OptimalClustersize/'

# Create a vector of file indices
nrclusters <- 5:30

# Loop through the file indices and read the tables, named x_1, 
#x being the nr of clusters, and _1 indicating it is the first replicate
for (i in nrclusters) {
  assign(paste0("x", i), read.table(paste0(folder, i, "_1.txt"), skip = 8))
}

clustertable <- read_tsv('OptimalClustersize/CollecteDataframe.txt')
clustertable <- clustertable[-nrow(clustertable), ]

results_df <- data.frame(cluster_number = integer(), mean_value = numeric())

for (i in 5:30) {
# Defining the heatmap to analyze and what cluster
heatmap_name <- paste0('x', i)
heatmap <- get(heatmap_name)
column <- paste('Cluster with', i)
clusterchosen <- clustertable[[column]]

# Adding cluster numbers in the individual dataframe and adding 1
clusterchosen <- sort(as.vector(clusterchosen))
heatmap$clusternr <- clusterchosen
allclusters <- 1:(nrow(table(heatmap$clusternr)))
heatmap$clusternr <- heatmap$clusternr+1

# customizing MAD function so the it is mean and not median
custom_mad <- function(x) {
  mad(x, center = mean(x))
}


mad_results <- c()
for (clusterx in allclusters) {
  df <- heatmap %>% filter(clusternr == clusterx)
  df <- df %>% select(-clusternr)
  mad_values <- apply(df, 2, custom_mad)
  mean_mad_value <- mean(mad_values)
  #print(mean(mad_values))
  mad_results <- c(mad_results, mean_mad_value)
}


mean(mad_results)

#with the zeros
#x = c(5, 10, 15, 20, 25, 30)
#y = c(0.4999286, 0.5065915, 0.531052, 0.5183138, 0.5489617, 0.5445573)
#plot(x,y, type = "l")


#zeros are removed
vector_no_zeros <- mad_results[mad_results != 0]
mean_value <- mean(vector_no_zeros)

# Append the results to the dataframe
results_df <- rbind(results_df, data.frame(cluster_number = i, mean_value = mean_value))

}

plot(results_df, type = 'l')


ggplot(data=results_df, aes(y=mean_value, x=cluster_number)) +
  labs(x = "k", y = "Mean cluster variation", title = "Mean cluster variation in DBA Heatmap") +
  geom_line()+
  geom_point() +
  geom_smooth() +
  theme(text=element_text(size=40)) +
  theme_minimal()


```



Loading the data that was pulled from the individual heatmaps
```{r}
# naming of file needs to be 3 digit name + rep1/rep2 + cluster with 2 digits ex 06 or 12
# remember to change the working directory/path/filenames so it fits the current analysis

# loading all heatmaps, here saved in a folder called heatmapdata2
files <- list.files(path = "heatmapdata2", full.names = TRUE)

# loading table from EaSeq where we have information of clusternumbers
clustertable <- read_tsv('OptimalClustersize/CollecteDataframe.txt')
clustertable <- clustertable[-nrow(clustertable), ]


# Iterating through all the heatmaps and saving them as matrices
for (file in files) {
currentdata <- read.table(file, skip = 8)

currentname <- str_sub(file, 14, 29)
#currentname <- gsub('0', '', currentname) maybe not needed, but would be nice for later maybe
currentcluster <- str_sub(file, 28, 29)
currentcluster <- gsub('^0', '', currentcluster)

# Defining what data to use and the amount of clusters, here named "Cluster with x", x being the current cluster
clustercolumn <- paste('Cluster', 'with', currentcluster, sep = " ")
heatmap <- currentdata
clusterchosen <- clustertable[[clustercolumn]]

# Adding cluster numbers in the individual dataframe and adding 1
clusterchosen <- sort(as.vector(clusterchosen))
heatmap$clusternr <- clusterchosen
allclusters <- 1:(nrow(table(heatmap$clusternr)))
heatmap$clusternr <- heatmap$clusternr+1


finalmatrix <- matrix(NA, nrow = length(allclusters), ncol = 199)

# Saving the average signal across each bin in each cluster + removing clusters that are between 1-5 insertions only
for (clusterx in allclusters) {
  value_counts <- heatmap %>% count(clusternr)
  values_to_remove <- value_counts %>% filter(n >= 1 & n <= 5) %>% pull(clusternr)
  heatmap <- heatmap %>% filter(!(clusternr %in% values_to_remove))
  df <- heatmap %>% filter(clusternr == clusterx) 
  df <- df %>% select(-clusternr) 
  mean_values <- apply(df, 2, mean)
  finalmatrix[clusterx,] <- mean_values
}

finalmatrix <- na.omit(finalmatrix)
currentmatrix <- paste(currentname)
assign(currentname, finalmatrix)

}

```

Second step, is calulating the absolute differences between strain/strain comparison,
together with the replicate/replicate comparison

```{r}
# calculating the difference between replicates and strain heatmaps, for each cluster
numbers <- c('05','06', '08', 10, 12, 14, 15, 16, 18, 20, 22, 24, 25, 26, 28, 30)

for (i in numbers) {
  rep1_var <- paste0("BL6rep1cluster", i)
  rep2_var <- paste0("BL6rep2cluster", i)
  result_var <- paste0("repmatrix", i)
  
  rep1 <- get(rep1_var)
  rep2 <- get(rep2_var)
  
  # Calculate the absolute difference and storing the results
  result <- abs(rep1 - rep2)
  assign(result_var, result)
}


for (i in numbers) {
  opp1_var <- paste0("BL6rep1cluster", i)
  opp2_var <- paste0("DBArep1cluster", i)
  result_var <- paste0("oppmatrix", i)
  
  opp1 <- get(opp1_var)
  opp2 <- get(opp2_var)
  
  # Calculate the absolute difference and storing the results
  result <- abs(opp1 - opp2)
  assign(result_var, result)
}

```


Lastly, generating different plots to give an idea of how these numbers actually look

```{r}
y <- c(mean(repmatrix06),mean(repmatrix08),mean(repmatrix10), mean(repmatrix12), mean(repmatrix14), mean(repmatrix16), mean(repmatrix18), mean(repmatrix20), mean(repmatrix22), mean(repmatrix24), mean(repmatrix26), mean(repmatrix28), mean(repmatrix30))
x <- c(6,8,10,12,14,16,18,20,22,24,26,28,30)

df <- data.frame(x, y)
ggplot(data=df, aes(x=x, y=y)) +
  geom_line()+
  ggtitle("Replicate average difference")+
  geom_point(size = 1) +
  theme_minimal()

y1 <- c(mean(oppmatrix06),mean(oppmatrix08),mean(oppmatrix10), mean(oppmatrix12), mean(oppmatrix14), mean(oppmatrix16), mean(oppmatrix18), mean(oppmatrix20), mean(oppmatrix22), mean(oppmatrix24), mean(oppmatrix26), mean(oppmatrix28), mean(oppmatrix30))

df1 <- data.frame(x, y1)
ggplot(data=df1, aes(x=x, y=y1)) +
  geom_line()+
  ggtitle("Strain/strain average difference")+
  geom_point(size = 1) +
  theme_minimal()


#combined strain/strain and replicates comparison
dfcombined <- data.frame(x, y, y1)
ggplot(dfcombined, aes(x=x)) + 
  geom_line(aes(y = y), color = "darkred") + 
  geom_line(aes(y = y1), color="steelblue")+
  theme_minimal()



x <- -99:99
y <- apply(repmatrix05, 2, mean)
y1 <- apply(repmatrix10, 2, mean)
y2 <- apply(repmatrix15, 2, mean)
y3 <- apply(repmatrix20, 2, mean)
y4 <- apply(repmatrix25, 2, mean)
y5 <- apply(repmatrix30, 2, mean)

# Create a data frame
df <- data.frame(
  "Position" = x,
  "clusters05" = y,
  "clusters10" = y1,
  "clusters15" = y2,
  "clusters20" = y3,
  "clusters25" = y4,
  "clusters30" = y5
)


longdf <- df %>% pivot_longer(!Position, names_to = "Number of clusters", values_to = "value")

#plots made with the long form
gfg_plot <- ggplot(longdf, aes(x = Position, y = value, color = `Number of clusters`)) +  
  geom_line() +
  theme_minimal()

# Display the plot
print(gfg_plot)


# Plot multiple lines using matplot
matplot(x, cbind(y, y1, y2, y3, y4, y5), type = "l", lty = 1, 
        col = c("#ADD8E6", "#87CEEB", "#6495ED", "#4682B4", "#4169E1", "#191970"), xlab = "Bin", 
        ylab = "Mean difference", main = "Average difference across bins for replicates")
legend("topright", legend = c("5 clusters", "10 clusters", "15 clusters", "20 clusters", "25 clusters", "30 clusters"), 
       col = c("#ADD8E6", "#87CEEB", "#6495ED", "#4682B4", "#4169E1", "#191970"), 
       lty = 1)



x <- -99:99
y <- apply(oppmatrix05, 2, mean)
y1 <- apply(oppmatrix10, 2, mean)
y2 <- apply(oppmatrix15, 2, mean)
y3 <- apply(oppmatrix20, 2, mean)
y4 <- apply(oppmatrix25, 2, mean)
y5 <- apply(oppmatrix30, 2, mean)

# Plot multiple lines using matplot
matplot(x, cbind(y, y1, y2, y3, y4, y5), type = "l", lty = 1, 
        col = c("#ADD8E6", "#87CEEB", "#6495ED", "#4682B4", "#4169E1", "#191970"), xlab = "Bin", 
        ylab = "Mean difference", main = "Average difference across bins for strain/strain difference")
legend("topright", legend = c("5 clusters", "10 clusters", "15 clusters", "20 clusters", "25 clusters", "30 clusters"), 
       col = c("#ADD8E6", "#87CEEB", "#6495ED", "#4682B4", "#4169E1", "#191970"), 
       lty = 1)
```



